Cascades, Product Conversion, Whatever

### 2012-12-12 paper.tex
Made small spelling/grammar changes

More specifically, we propose to research the impact of having
strategic users: how much different are results if we assume users are
strategic instead of myopic. To answer this question, we will use a
model introduced in \cite{Chierichetti12} and described in
Section~\ref{prob_statement}. Our contribution is to fully
characterize the behavior of this model on a specific class of graphs,
the 2-blockmodels, in hopes of learning more about the general
case. Blockmodels have been studied extensively in the past
\cite{Wang87, Snijders97} as a natural framing of networks in which we
can think of a node having a type which influences its
characteristics. For example, a blockmodel describing a social network
could have types for each grade, and students with each type are more
likely to connect to their own grade, and unlikely to connect to
grades very far away.

The analysis of strategic behavior on networks is not a new
concept. Kearns et. al.  introduced a convenient method for expressing
games of this sort with their seminal work on graphical games
\cite{Kearns01}. But, to the extent of our knowledge, nobody has
attempted to characterize full graphical game models for cascade
spreading.  This is what we do below.

@@

We keep the same scheduler behavior as in the previous paper. The
schedulers utility is based off of the number of nodes that choose Yes
at the end of the game. We assume the scheduler can change its plan
after it sees what a node chooses. We also assume it has full
knowledge of the nodes strategies, as well as knowledge of the full
graph, and any game parameters. The only aspect the scheduler does not
know is the individual nodes types.

The nodes utility is the sum of all of their neighbors that match
their choice plus some constant ($\pi$) if their choice matches their
type. Strategic nodes make decisions to maximize their expected
profit. We assume strategic nodes know the schedulers decision making
(or utility function), the graph structure, and the probability that a
node is assigned Yes type ($p$), but not any other nodes type. To keep
the model interesting, we restrict $p < 0.5$. Similarly myopic nodes
only try to maximize their current utility, ignoring any nodes that
has not made a decision.

We consider games where either all nodes in the graph are strategic or
myopic, and want to see what graphs / game parameters contribute to
significant differences between the expected number of Yesses when
nodes are strategic versus myopic.

We formally specify the model as follows
\begin{align*}
  & G(V,E) && \text{social network} \\
  & |V| = n \\
  & c_i \in \{Y,N\} && \text{choice of node $i$} \\
  & t_i \in \{Y, N\} && \text{type of node $i$} \\
  & p(t_i = Y) = p < 0.5 \\
  & u_i(\mathbf c) = \pi \mathbb I[t_i = c_i] + \sum_{j \in N(i)}
  \mathbb I[c_j = c_i] && \text{node utility} \\
  & sched(\mathbf c) && \text{schedule function}
\end{align*}
where $\mathbb I$ is the indicator function, and $N$ is the neighbor
function. The schedule function returns the next node to pick given
the choices that have been made so far.

@@

Is seems likely that computing the strategic choices and payouts is a
hard problem, but we will show an algorithm that can compute an
optimal online schedule and calculate $\mathbb E[\#Y|stragetic]$ in
polynomial time for arbitrary block models.

@@

The two tables recursively reference each other to calculate their
entries, where each entry is the choice, and the expected number of
Yesses at the end of the game in each block conditioned on that
choice. If an entry is already calculated, then the lookup is constant
time, otherwise the lookup may cause a cascade of calculations, which
is bounded by the polynomial time to compute both tables.\footnote{It
  is possible that instead of looking these up on the fly, he table
  could be precomputed to speed up the complexity slightly, but we
  believe that this speedup would be very slight, and would make the
  algorithm much more complicated}

@@

To determine the optimal block choice for the scheduler, the decision
function only needs to know, ho many nodes from each block have
already gone, and what proportion (or number) of the nodes have chosen
Yes. Because nodes within a block are exchangeable, this information
completely characterizes the state of the game at the time of
scheduler choice.

@@

A node needs to check its own expected final utility given it chooses
Yes or No, and choose the choice that maximizes it. The node can query
the scheduler function to determine the expected number of Yesses in
each block conditioned on the node choosing Yes or No. The node then
trivially calculates its expected utility for each choice, and picks
the maximum.

@@

Lookups are constant time, so the total time complexity is going to
be bounded by the number of entries in the table times the number of
table lookups it takes to compute one entry in the table.

The size of the node table is trivial to upper bound. There are $2k+2$
dimensions: the node's type (2 values), the node's block ($k$ values),
$k$ dimensions for the number of nodes in each block that have gone
before ($\le s_i + 1 | i=1..k$), and $k$ dimensions for the number of
nodes in each block that have gone before and selected yes ($\le s_i +
1 | i=1..k$). This leads to the following statement,
\begin{align*}
  NodeTableSize &= o\left( 2k \prod_{i=1}^k s_i+1 \prod_{i=1}^k s_i+1
  \right) \\
  &= o\left( k \prod_{i=1}^k s_i^2 \right)
  \intertext{If we further assume that $\ell \le k$ of
    the blocks are size order $n$ and the rest are constant size, then the
    size of the table reduces to}
  NodeTableSize &= o\left( kn^{2\ell} \right)
  \intertext{Each node table entry
    computation takes 2 calls to the sched function, and then does
    computation of the expected number of Yeses which is order $k$,
    making the total complexity for the node table given the sched
    table}
  NodeTableComplexity &= o\left( k^2 \prod_{i=1}^k s_i^2 \right) \\
  &= o\left( k^2 n^{2\ell} \right)
  \intertext{The size of the scheduler table is also trivial to upper
    bound. There are $2k$ dimensions similar to the node table. By
    similar reasoning}
  SchedTableSize &= o\left( \prod_{i=1}^k s_i+1 \prod_{i=1}^k s_i+1
  \right) \\
  &= o\left( \prod_{i=1}^k s_i^2 \right)
  \intertext{If we make the same assumption about block sizes this
    reduces to}
  SchedTableSize &= o \left( n^{2\ell} \right)
  \intertext{The complexity for each entry in the scheduler table is
    linear in $k$, because it makes $2k$ lookups to the node
    table. Therefore the total complexity for constructing the
    scheduler table is bounded by}
  SchedTableComplexity &= o\left( k \prod_{i=1}^k s_i^2 \right) \\
  &= o \left( k n^{2\ell} \right)
  \intertext{This leaves us with final time complexity of}
  RunningTime &= o\left( k^2 \prod_{i=1}^k s_i^2 \right) \\
  &= o\left( k^2 n^{2\ell} \right)
\end{align*}

@@

We have only shown a polynomial time algorithm for graphs where the
number of blocks is constant with respect to $N$. A further and
important step would be to come up with an algorithm for general
graphs, or to reduce the problem to another complexity class like PPAD
or NP.

>>>>

More specifically, we propose to research the impact of having
strategic users: how different are results if we assume users are
strategic instead of myopic? To answer this question, we will use a
model introduced in \cite{Chierichetti12} and described in
Section~\ref{prob_statement}. Our contribution is to begin to
characterize the behavior of this model on a specific class of graphs,
the blockmodels, in hopes of learning more about the general
case. Blockmodels have been studied extensively in the past
\cite{Wang87, Snijders97} as a natural framing of networks in which we
can think of a node having a type which influences its
characteristics. For example, a blockmodel describing a social network
could have types for each grade, and students with each type are more
likely to connect to their own grade, and unlikely to connect to
grades very far away.

The analysis of strategic behavior on networks is not a new
concept. Kearns et. al.  introduced a convenient method for expressing
games of this sort with their seminal work on graphical games
\cite{Kearns01}. But, to the best of our knowledge, nobody has
attempted to solve graphical game models for cascade
spreading.  This is what we do below.

@@

We keep the same scheduler behavior as in the previous paper. The
scheduler's utility is based off of the number of nodes that choose Yes
at the end of the game. We assume the scheduler can change its plan
after it sees what a node chooses. We also assume it has full
knowledge of the nodes' strategies, as well as knowledge of the full
graph, and any game parameters. The only aspect the scheduler does not
know is the individual node types.

The nodes' utility is the sum of all of their neighbors that match
their choice plus some constant ($\pi$) if their choice matches their
type. Strategic nodes make decisions to maximize their expected
profit. We assume strategic nodes know the schedulers decision making
(or utility function), the graph structure, and the probability that a
node is assigned Yes type ($p$), but not any other nodes type. To keep
the model interesting, we restrict $p < 0.5$. Conversely, myopic nodes
only try to maximize their current utility, ignoring any nodes that
have not made a decision.

We consider games where either all nodes in the graph are strategic or
all are myopic, and want to see what graphs / game parameters contribute to
significant differences between the expected number of Yesses when
nodes are strategic versus myopic.

We formally specify the model as follows
\begin{align*}
  & G(V,E) && \text{social network} \\
  & |V| = n \\
  & c_i \in \{Y,N\} && \text{choice of node $i$} \\
  & t_i \in \{Y, N\} && \text{type of node $i$} \\
  & p(t_i = Y) = p < 0.5 \\
  & u_i(\mathbf c) = \pi \mathbb I[t_i = c_i] + \sum_{j \in N(i)}
  \mathbb I[c_j = c_i] && \text{node utility} \\
  & sched(\mathbf c) && \text{schedule function}
\end{align*}
where $\mathbb I$ is the indicator function, and $N$ is the neighborhood
function. The schedule function returns the next node to pick given
the choices that have been made so far.

@@

Is seems likely that computing the strategic choices and payouts is a
hard problem, but we will show an algorithm that can compute an
optimal online schedule and calculate $\mathbb E[\#Y|stragetic]$ for arbitrary
block models in time polynomial in the number of blocks.

@@

The two tables recursively reference each other to calculate their
entries, where each entry is the choice, and the expected number of
Yesses at the end of the game in each block conditioned on that
choice. If an entry is already calculated, then the lookup is constant
time, otherwise the lookup may cause a cascade of calculations, which
is bounded by the polynomial time to compute both tables.\footnote{It
  is possible that instead of looking these up on the fly, the table
  could be precomputed to speed up the complexity slightly, but we
  believe that this speedup would be very slight, and would make the
  algorithm much more complicated}

@@

To determine the optimal block choice for the scheduler, the decision
function only needs to know how many nodes from each block have
already gone and what proportion (or number) of the nodes have chosen
Yes. Because nodes within a block are exchangeable, this information
completely characterizes the state of the game at the time of
scheduler choice.

@@

A node needs to check its own expected final utility given it chooses
Yes or No, and choose the choice that maximizes it. The node can query
the scheduler function to determine the expected number of Yesses in
each block conditioned on the node choosing Yes or No. The node then
trivially calculates its expected utility for each choice and picks
the maximum.

@@

Lookups are constant time, so the total time complexity is 
 bounded by the number of entries in the table times the number of
table lookups it takes to compute one entry in the table.

The size of the node table is trivial to upper bound. There are $2k+2$
dimensions: the node's type (2 values), the node's block ($k$ values),
$k$ dimensions for the number of nodes in each block that have gone
before ($\le s_i + 1 | i=1..k$), and $k$ dimensions for the number of
nodes in each block that have gone before and selected yes ($\le s_i +
1 | i=1..k$). This leads to the following statement,
\begin{align*}
  NodeTableSize &= O\left( 2k \prod_{i=1}^k s_i+1 \prod_{i=1}^k s_i+1
  \right) \\
  &= O\left( k \prod_{i=1}^k s_i^2 \right)
  \intertext{If we further assume that $\ell \le k$ of
    the blocks are size order $n$ and the rest are constant size, then the
    size of the table reduces to}
  NodeTableSize &= O\left( kn^{2\ell} \right)
  \intertext{Each node table entry
    computation takes 2 calls to the sched function, and then does
    computation of the expected number of Yeses which is order $k$,
    making the total complexity for the node table given the sched
    table}
  NodeTableComplexity &= O\left( k^2 \prod_{i=1}^k s_i^2 \right) \\
  &= O\left( k^2 n^{2\ell} \right)
  \intertext{The size of the scheduler table is also trivial to upper
    bound. There are $2k$ dimensions similar to the node table. By
    similar reasoning}
  SchedTableSize &= O\left( \prod_{i=1}^k s_i+1 \prod_{i=1}^k s_i+1
  \right) \\
  &= O\left( \prod_{i=1}^k s_i^2 \right)
  \intertext{If we make the same assumption about block sizes this
    reduces to}
  SchedTableSize &= O \left( n^{2\ell} \right)
  \intertext{The complexity for each entry in the scheduler table is
    linear in $k$, because it makes $2k$ lookups to the node
    table. Therefore the total complexity for constructing the
    scheduler table is bounded by}
  SchedTableComplexity &= O\left( k \prod_{i=1}^k s_i^2 \right) \\
  &= O \left( k n^{2\ell} \right)
  \intertext{This leaves us with final time complexity of}
  RunningTime &= O\left( k^2 \prod_{i=1}^k s_i^2 \right) \\
  &= O\left( k^2 n^{2\ell} \right)
\end{align*}

@@

We have only shown a polynomial time algorithm for graphs where the
number of blocks is constant with respect to $n$. A further and
important step would be to come up with an algorithm for general
graphs, or to reduce the problem to another complexity class like PPAD
or NP.
